---
title: "Teaching"
---
```{css, echo = FALSE}

d-article h1 {
  font-size: 28px;
  padding-bottom: .5rem;
}

d-article h3 {
margin-top: .25rem;
  margin-bottom: 6px;
  font-size: 14px;
color: rgba(42, 25, 79, .6)!important;
}

d-article > p {
  padding-bottom: 20px;
}

.l-body {
  font-size: 1rem;
  padding-top: 4px;
}

.l-body p {
  margin-bottom: 6px
}

h2 a:hover {
    text-decoration: none;
}

```

# Harvard TH Chan School of Public Health

## [EPI 207: Advanced Epidemiologic Methods](https://courses.my.harvard.edu/psp/courses/EMPLOYEE/EMPL/h/?tab=HU_CLASS_SEARCH&SearchReqJSON=%7B%22PageNumber%22%3A1%2C%22PageSize%22%3A%22%22%2C%22SortOrder%22%3A%5B%22IS_SCL_SUBJ_CAT%22%5D%2C%22Facets%22%3A%5B%5D%2C%22Category%22%3A%22HU_SCL_SCHEDULED_BRACKETED_COURSES%22%2C%22SearchPropertiesInResults%22%3Atrue%2C%22FacetsInResults%22%3Atrue%2C%22SaveRecent%22%3Atrue%2C%22TopN%22%3A%22%22%2C%22SearchText%22%3A%22EPI%20207%20Robins%22%2C%22DeepLink%22%3Afalse%7D)

### (Head) Teaching fellow (fall 2018, fall 2019, fall 2020)

Causal inference for time-varying exposures: g-formula, marginal structural models, g-estimation of structural nested models; static and dynamic treatment regimes.
  
## [BST 256: Theories and Methods for Causality I](https://courses.my.harvard.edu/psp/courses/EMPLOYEE/EMPL/h/?tab=HU_CLASS_SEARCH&SearchReqJSON=%7B%22ExcludeBracketed%22%3Atrue%2C%22PageNumber%22%3A1%2C%22PageSize%22%3A%22%22%2C%22SortOrder%22%3A%5B%22IS_SCL_SUBJ_CAT%22%5D%2C%22Facets%22%3A%5B%5D%2C%22Category%22%3A%22HU_SCL_SCHEDULED_BRACKETED_COURSES%22%2C%22SearchPropertiesInResults%22%3Atrue%2C%22FacetsInResults%22%3Atrue%2C%22SaveRecent%22%3Atrue%2C%22TopN%22%3A%22%22%2C%22SearchText%22%3A%22BST%20256%22%2C%22DeepLink%22%3Afalse%7D)
  
### Teaching fellow (fall 2020)

Structural causal models, graphical models, algorithms for identification.
  

## [Introduction to R](https://intro-to-R-2020.louisahsmith.com)
### Instructor (fall 2019, summer 2020)
```{r, echo = FALSE}
distilltools::icon_link(
  icon = "chalkboard-teacher",
  text = "Course materials",
  url = "https://intro-to-R-2020.louisahsmith.com"
)
```

Developed and taught short course to introduce R to graduate students. Topics included creating figures and tables, cleaning data, writing functions.
  
## [Introduction to Epidemiology](https://www.hsph.harvard.edu/biostatistics/diversity/summer-program/)
  
### Co-instructor (summer 2019, summer 2020)

Epidemiology course for participants in the Harvard Summer Program in Biostatistics & Computational Biology.

## [PHS 2000: Quantitative Research Methods in Population Health Sciences](https://courses.my.harvard.edu/psp/courses/EMPLOYEE/EMPL/h/?tab=HU_CLASS_SEARCH&SearchReqJSON=%7B%22PageNumber%22%3A1%2C%22PageSize%22%3A%22%22%2C%22SortOrder%22%3A%5B%22IS_SCL_SUBJ_CAT%22%5D%2C%22Facets%22%3A%5B%5D%2C%22Category%22%3A%22HU_SCL_SCHEDULED_BRACKETED_COURSES%22%2C%22SearchPropertiesInResults%22%3Atrue%2C%22FacetsInResults%22%3Atrue%2C%22SaveRecent%22%3Atrue%2C%22TopN%22%3A%22%22%2C%22SearchText%22%3A%22PHS%202000%22%2C%22DeepLink%22%3Afalse%7D)
### Teaching fellow (fall 2017, spring 2018, fall 2019)
```{r, echo = FALSE}
distilltools::icon_link(
  icon = "chalkboard-teacher",
  text = "Summer prep site",
  url = "https://phs2000-prep.hmdc.harvard.edu"
)
```
Regression models, sampling, longitudinal and multilevel analysis, time-varying confounding, mediation and interaction, econometric methods, missing data.
  
# University of California, Berkeley

## [PH 150A: Introduction to Epidemiology and Human Disease](http://guide.berkeley.edu/courses/pb_hlth/)
### Graduate student instructor (spring 2016)

Basic epidemiologic methods, overview of epidemiology of diseases/conditions of public health importance for undergraduates.


## [PH 250A: Epidemiologic Methods](http://guide.berkeley.edu/courses/pb_hlth/)
### Graduate student instructor (fall 2015)

Study design, sampling, data collection, analysis, inference for MPH students.
